Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-10-10 methods_paper/fabl-paper.tex
comments on section 1.

\noindent%
{\it Keywords:} record linkage, bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo, parallel/distributed computing

@@

	I have some proposed revisions for the first part of the paper. In the introduction, I reordered things and changed some topic sentences to highlight that the main contribution here is speed and scale.

	Also, in the prior version, we didn't get to any of the proposed material until Section 4.3, which did not seem reasonable to me.
	I know you recommended splitting the review of the literature between the Introduction and Section 2 on bipartite matching, but since we needed the mathemetical overview of FS and Sadinle, I decided to move all of the literature review into the introduction. I've seen this done in many papers.

	Almost all of the content is still there, but it's moved around in a way that makes more sense to me. Let me know your thoughts! (But please avoid hard edits until we talk on Monday!)

	I have some additional edits in the later part of the paper about the partial Bayes estimate, computational complexity, and added detail for the case study. All of that is in blue as well.

	---------------------------------------------

	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}. This is an increasingly important task in "data cleaning," and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others. In this paper, we consider bipartite record linkage, which is the merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.

@@

	In this paper, we propose \emph{fast beta linkage} (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a "many-to-one" model for record linkage. This allows us to employ hashing techniques that hasten calculations and reduce computational complexity, complete pairwise record comparisons over large datasets through parallel computing, and reduce memory costs through a new technique called \emph{storage efficient indexing}. Throughout, we argue that even in cases where a bipartite matching is desired, this relaxed model (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed. To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl},
	pronounced ``fable'')

	In what follows, we first review more thoroughly the Fellegi Sunter and Beta Record Linkage methods. Then, we provide the model specification under \texttt{fabl} and derive the Gibbs sampler for posterior inference. Next, we introduce the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake. Lastly, we use simulation studies and a case study of homicides from the El Salvadoran Civil War to show that our method has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.

@@

Consider two databases $\bm{X}_1$ and $\bm{X}_2$ with respective sizes $n_1$ and $n_2$. Without loss of generality, denote files suchs that $n_1 \geq n_2$. In the context of bipartite matching, we assume that there are duplications across, but not within, each database. Under this framework, the set of matches across datasets can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1 \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0 \quad \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become costly for large linkage tasks or for long posterior samplers. More compactly, bipartite matching can also be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j \quad  \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
	\end{cases}
\end{align}
Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot)$ is the indicator function.

>>>>

\noindent%
{\it Keywords:} bipartite record linkage, Bayesian methods, Gibbs sampling, hashing techniques, Markov chain Monte carlo, parallel/distributed computing

@@

%	I have some proposed revisions for the first part of the paper. In the introduction, I reordered things and changed some topic sentences to highlight that the main contribution here is speed and scale. 
%	
%	Also, in the prior version, we didn't get to any of the proposed material until Section 4.3, which did not seem reasonable to me.
%	I know you recommended splitting the review of the literature between the Introduction and Section 2 on bipartite matching, but since we needed the mathemetical overview of FS and Sadinle, I decided to move all of the literature review into the introduction. I've seen this done in many papers. 
%	
%	Almost all of the content is still there, but it's moved around in a way that makes more sense to me. Let me know your thoughts! (But please avoid hard edits until we talk on Monday!)
%	
%	I have some additional edits in the later part of the paper about the partial Bayes estimate, computational complexity, and added detail for the case study. All of that is in blue as well.

	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}. This is an increasingly important task in ``data cleaning," and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others. In this paper, we consider bipartite record linkage, which is merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.

@@

	In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a ``many-to-one" model for record linkage. This allows us to employ hashing techniques that hasten calculations and reduce computational costs. Furthermore, we can compute the pairwise record comparisons over large datasets via parallel computing and reduce memory costs through our proposed method of \emph{storage efficient indexing}. Throughout, we argue that even in cases where a bipartite matching is desired, \textcolor{red}{our proposed approach and relaxation} (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed. We illustrate our results via simulation studies and a real case study of homicides from the El Salvadoran Civil War.

\textcolor{red}{BK: I would layout all the sections, section by section once they have been set for the reader and I can help you do this.}	
	% I think this is clear and has already been said. 
%	To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl}, pronounced ``fable'')

%	In what follows, we first review more thoroughly the Fellegi Sunter and Beta Record Linkage methods. Then, we provide the model specification under \texttt{fabl} and derive the Gibbs sampler for posterior inference. Next, we introduce the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake. Lastly, we use simulation studies and a case study of homicides from the El Salvadoran Civil War to show that our method has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.

@@

Consider two databases $\bm{X}_1$ and $\bm{X}_2$ with respective sizes $n_1$ and $n_2$. Without loss of generality, denote files such that $n_1 \geq n_2$. In the context of bipartite matching, we assume that there are duplications across, but not within, each database. Under this framework, the set of matches across datasets can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1 \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0 \quad \text{otherwise}.\\
	\end{cases}
\end{align}
Though intuitive, this sparse matrix representation can become costly for large linkage tasks or for long posterior samplers. More compactly, bipartite matching can also be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in database $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i \quad \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j \quad  \quad \text{if records}\;  j \in \bm{X}_2 \; \text{does not have a match in database}\; \bm{X}_1. \\
	\end{cases}
\end{align}
Depending on which representation is the most convenient, we can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot)$ is the indicator function.
